---
title: "`r glue('{cur_event$SHORT.CODE} {cur_event$EDITION} summary')`"
subtitle: "`r glue('{cur_event$START.DATE} to {cur_event$END.DATE}')`"
author: "Bird Count India"
date: "`r Sys.Date()`"
output: 
  bookdown::word_document2:
  fig_caption: yes
  reference_docx: word_template.docx
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)


library(tidyverse)
library(lubridate)
library(glue)
library(writexl)
library(readxl)
library(flextable)
library(rebird)
# library(skimmr)


ebd_tax <- read_csv("eBirdTaxonomy.csv")

source("scripts/functions_anal.R")
source("token.R")
source("https://raw.githubusercontent.com/birdcountindia/bci-functions/main/01_functions/summaries.R")
source("https://raw.githubusercontent.com/birdcountindia/bci-functions/main/01_functions/mapping.R")

# get required dates
get_dates(seq(cur_event$START.DATE, cur_event$END.DATE, by = "days"))

```

```{r data-prep, include=FALSE}

# preparing data 
# filtering for loc & dates

# get required admin units
cur_unit_code <- if (cur_event$SHORT.CODE == "TEBC"){
  "IN-NL"
} else if (cur_event$SHORT.CODE == "SABC"){
  "IN"
} else if (cur_event$SHORT.CODE == "PBC"){
  "IN-TN"
} else if (cur_event$SHORT.CODE == "BBC"){
  "IN-AS"
}

cur_region <- get_admin_codes(hi_arch = FALSE, unit_code = cur_unit_code)
cur_region_children <- get_admin_codes(cur_region, hi_arch = TRUE) %>% 
  setdiff(cur_region)

which_admin <- if (str_count(cur_region, "-") == 0) "country" else {
  if (str_count(cur_region, "-") == 1) "state" else "district"
}

data0 <- data %>% 
  filter(OBSERVATION.DATE %in% dates_cur) %>% 
  {if (which_admin == "state") {
    filter(., STATE.CODE == cur_region)
  } else if (which_admin == "district") {
    filter(., COUNTY.CODE == cur_region)
  } else {
    .
  }}

```

```{r data-processing, include=FALSE}

# species lists & participation summaries
spec_list_main <- gen_spec_list_nonAPI(data0, cur_region, dates_cur)
part_summ_main <- gen_part_summ_nonAPI(data0, cur_region, dates_cur)

spec_list_sub <- gen_spec_list_nonAPI(data0, cur_region_children, dates_cur)
part_summ_sub <- gen_part_summ_nonAPI(data0, cur_region_children, dates_cur) 

# notable species (SoIB)
spec_list_notable <- map_to_soib(spec_list_main) %>% 
  filter(SOIBv2.Priority.Status == "High") %>% 
  slice_head(n = 10)


# functions to create statements from summaries
create_statement <- function(var) {
  
  if (var == "observers") {
    
    where <- get_admin_names(cur_region) %>% 
      filter(REGION == cur_region) %>% 
      pull(REGION.NAME)
    
    if (length(dates_cur) == 1) {
      
      number <- part_summ_main$OBSERVERS %>% as.integer()
      string <- glue("**{number} birdwatchers** from {where} came together")
      
      return(string)
      
    } else if (length(dates_cur) > 1) {
      
      info <- part_summ_main %>% 
        filter(TOTAL == "OBSERVERS") %>% 
        dplyr::select(starts_with("DAY")) %>% 
        pivot_longer(cols = everything(), names_to = "DAY.NO", values_to = "OBSERVERS") %>% 
        mutate(MAX = max(OBSERVERS), 
               MAX.WHEN = case_when(OBSERVERS == MAX ~ DAY.NO, 
                                    TRUE ~ NA),
               MIN = min(OBSERVERS),
               NO.DAYS = n_distinct(DAY.NO)) %>% 
        distinct(MAX, MAX.WHEN, MIN, NO.DAYS) %>% 
        filter(!is.na(MAX.WHEN))
      
      string <- glue("**Everyday, at least {info$MIN} birdwatchers** from {where} came together for the {info$NO.DAYS} days ({info$MAX} birdwatchers on Day {info$MAX.WHEN %>% str_remove('DAY')}!),")
      
      return(string)
      
    }
    
  } else if (var == "species") {
    
    if (length(dates_cur) == 1) {
      number <- part_summ_main$SPECIES %>% as.integer()
    } else if (length(dates_cur) > 1) {
      number <- part_summ_main %>% 
        filter(TOTAL == "SPECIES") %>% 
        pull(ALL.DAYS) %>% 
        as.integer()
    }
    
    string <- glue("**{scales::label_comma()(number)} species**")
    return(string)
    
  } else if (var == "checklists") {
    
    if (length(dates_cur) == 1) {
      number <- part_summ_main$CHECKLISTS %>% as.integer()
    } else if (length(dates_cur) > 1) {
      number <- part_summ_main %>% 
        filter(TOTAL == "CHECKLISTS") %>% 
        pull(ALL.DAYS) %>% 
        as.integer()
    }
    
    string <- glue("**{scales::label_comma()(number)} checklists**")
    return(string)
    
  } else if (var == "subregions") {
    
    info <- part_summ_sub %>% 
      {if (length(dates_cur) == 1) {
        filter(., if_any(c("OBSERVERS", "CHECKLISTS", "SPECIES"), ~ .x > 0))
      } else if (length(dates_cur) > 1) {
        filter(., if_any(starts_with("DAY"), ~ .x > 0))
      }} %>% 
      reframe(NO.REG = n_distinct(REGION),
              WHICH.REG = case_when(str_count(REGION, "-") == 1 ~ "states/union territories",
                                    str_count(REGION, "-") == 2 ~ "districts",
                                    TRUE ~ "UNKNOWN REGION")) %>% 
      distinct(NO.REG, WHICH.REG)
    
    string <- glue("**{info$NO.REG} {info$WHICH.REG}**")
    return(string)
    
  } else if (var == "notable") {
    
    if (length(spec_list_notable) > 0) {
      
      info <- spec_list_notable %>% 
        pull(COMMON.NAME) %>% 
        str_flatten_comma()
      
      string <- glue("Among the species reported were the following classed as of High Priority in the [SoIB 2023 report](https://stateofindiasbirds.in/#soib_report): {info}.")
      
    } else {
      string <- ""
    }
    
    return(string)
    
  }
  
}

```

```{r statements, include=FALSE}

how_many_people <- create_statement("observers")
how_many_species <- create_statement("species")
how_many_lists <- create_statement("checklists")
how_many_subregions <- create_statement("subregions")

what_subregions <- how_many_subregions %>% 
  str_extract_all("[:letter:]", simplify = TRUE) %>% 
  str_flatten()

which_notable_species <- create_statement("notable")

```

<br>
  
[Paragraph about event intro, coincidence with global event if so]

`r how_many_people` and documented `r how_many_species` in `r how_many_lists`! Participants represented `r how_many_subregions`. `r which_notable_species`

[Placeholder for nice ML image]

Thanks to all birdwatchers and nature enthusiasts from across the country who made this event a success. A special shout out to all those who took this opportunity to conduct public bird walks and introduce new people to the wonders of birds and nature!
  
```{r table1, tab.cap=glue("Total number of lists uploaded from {what_subregions}"), out.width="100%"}

tab1 <- part_summ_sub %>% 
  filter(TOTAL == "CHECKLISTS") %>% 
  dplyr::select(-TOTAL) %>% 
  arrange(desc(as.numeric(ALL.DAYS)))

write_csv(tab1, glue("{cur_outpath}summary_tab1.csv"))

tab1 %>% 
  flextable() %>% 
  set_table_properties(layout = "autofit") %>% 
  align(align = "center", part = "all") %>%
  align(align = "left", j = c("REGION", "REGION.NAME"), part = "all") %>%
  line_spacing(space = 1, part = "all") %>% 
  padding(padding = 1.5, part = "body") %>%
  fontsize(size = 8, part = "all") %>% 
  flextable::font(fontname = "Times New Roman", part = "all")

```

<br>
  
```{r table2, tab.cap=glue("Total number of species reported from {what_subregions}"), out.width="100%"}

tab2 <- part_summ_sub %>% 
  filter(TOTAL == "SPECIES") %>% 
  dplyr::select(-TOTAL) %>% 
  arrange(desc(as.numeric(ALL.DAYS)))

write_csv(tab2, glue("{cur_outpath}summary_tab2.csv"))

tab2 %>% 
  flextable() %>% 
  set_table_properties(layout = "autofit") %>% 
  align(align = "center", part = "all") %>%
  align(align = "left", j = c("REGION", "REGION.NAME"), part = "all") %>%
  line_spacing(space = 1, part = "all") %>% 
  padding(padding = 1.5, part = "body") %>%
  fontsize(size = 8, part = "all") %>% 
  flextable::font(fontname = "Times New Roman", part = "all")

```

<br>
  
[Placeholder for birding image]

Thank you to everyone for interacting with the public and students, organising bird walks and talks, and promoting the joy of birdwatching!
  